\documentclass{./../../Latex/handout}
\begin{document}
\thispagestyle{plain}
\myheader{Handout for Lecture 19 \\ \normalsize{Categorical Variables \& Interaction Terms}}
\rhead{Class Handout for Lecture 19}
\vspace{-1em}

\begin{enumerate}
\item Consider the following regression model:
$$ Y = \beta_0 + \beta_1 D + \beta_2 X + \beta_3  D\cdot X + u $$
Here, \( X \) represents a continuous variable, and \( D \) is a dummy variable that takes values 1 or 0. Assume that both $X$ and $D$ are exogenous. Write down the expressions for the following expectations.
$$ E(Y|D=1, X) =  \hspace{8cm} $$
$$ E(Y|D=0, X) = \hspace{8cm}  $$ 

What is the impact of changing $D$ from 1 to 0 on $Y$? Does this impact vary by $X$?

\vspace{3cm}
\item Consider the following regression model:$$ wages = \beta_0 + \beta_1 Female + \beta_2 Hispanic + \beta_3 Female \times Hispanic +  u  $$
Here, \(Female\) is a dummy variable assigned the value of 1 if an individual's gender is female and 0 if not. Similarly, \(Hispanic\) is a dummy variable that is set to 1 if an individual's ethnicity is Hispanic and 0 otherwise. The regression output for this model is given below. Answer the following questions.

\begin{table}
\centering
	\input{./../../output/gender_race_wage.tex}	
\end{table}

\newpage
\begin{enumerate}
	\item What is the average wage income for non-Hispanic males in this sample?
	\vspace{2cm}
	\item What is the average wage income for Hispanic males in this sample? \vspace{2cm}
	 \item What is the average wage income for non-Hispanic females in this sample?\vspace{2cm}
	\item What is the average wage income for Hispanic females in this sample? \vspace{2cm}
	\item How do we interpret the coefficient on the interaction between $Hispanic$ and $Female$?
\end{enumerate}
\end{enumerate}

\end{document}