\documentclass{./../../Latex/handout}
\begin{document}
\thispagestyle{plain}
\myheader{Handout for Lecture 19 \\ \normalsize{Categorical Variables \& Interaction Terms}}
\rhead{Class Handout for Lecture 19}
\vspace{-1em}

\begin{enumerate}
\item Consider the following regression model:
$$ Y = \beta_0 + \beta_1 D + \beta_2 X + \beta_3  D\cdot X + u $$
Here, \( X \) represents a continuous variable, and \( D \) is a dummy variable that takes values 1 or 0. Assume that both $X$ and $D$ are exogenous. Write down the expressions for the following expectations.
$$ E(Y|D=1, X) =  (\beta_0 + \beta_1) + (\beta_2+\beta_3) X \hspace{4cm} $$
$$ E(Y|D=0, X) = \beta_0 + \beta_2  X \hspace{6.5cm}  $$ 

What is the impact of changing $D$ from 1 to 0 on $Y$? Does this impact vary by $X$?
$$ E(Y|D=1, X)-E(Y|D=0, X) = \beta_1 + \beta_3 X $$
Yes, varies by $X$. \\~\\
\item Consider the following regression model:$$ wages = \beta_0 + \beta_1 Female + \beta_2 Hispanic + \beta_3 Female \times Hispanic +  u  $$
Here, \(Female\) is a dummy variable assigned the value of 1 if an individual's gender is female and 0 if not. Similarly, \(Hispanic\) is a dummy variable that is set to 1 if an individual's ethnicity is Hispanic and 0 otherwise. The regression output for this model is given below. Answer the following questions.

\begin{table}
\centering
	\input{./../../output/gender_race_wage.tex}	
\end{table}

\newpage
\begin{enumerate}
	\item What is the average wage income for non-Hispanic males in this sample?
	$$ E(wages |Female = 0, Hispanic = 0)  = \text{\$70,179.09} \hspace{3cm} $$
	\item What is the average wage income for Hispanic males in this sample? 	
	$$ E(wages |Female = 0, Hispanic = 1)  = \text{70,179.09-19,367.71} = \text{\$50,811.38}  \hspace{1cm} $$
	 \item What is the average wage income for non-Hispanic females in this sample?	$$ E(wages |Female = 1, Hispanic = 0)  = \text{70,179.09-16,046.81} = \text{\$54,132.28}  \hspace{1cm} $$
	\item What is the average wage income for Hispanic females in this sample?
	\begin{align*}
		E(wages |Female = 1, Hispanic = 1)  &= \text{70,179.09-19,367.71-16,046.81+8,163.75} \\ &= \text{\$42,928.32} 
	\end{align*} 	
	\newpage
	\item How do we interpret the coefficient on the interaction between $Hispanic$ and $Female$?
	
	It is the difference between the gender-wage gap for Hispanic vs non-Hispanic individuals. 
\begin{align*}
 E(wages |Female = 0, Hispanic = 1)-& E(wages |Female = 1, Hispanic = 1) \\
 & = \text{50811.38- 42928.32} = \text{7883.06} \\
 E(wages |Female = 0, Hispanic = 0)-& E(wages |Female = 1, Hispanic = 0)  \\
  & = \text{70179.09- 54132.28} = \text{16046.8}  
\end{align*} 	
Difference between the two: $\text{16046.8-7883.06}=\text{\$8,163.73}$

Alternatively, can interpret $\beta_3$ as the difference in the impact of being hispanic by gender. 
\end{enumerate}
\end{enumerate}

\end{document}