\documentclass{./../../Latex/handout}
\begin{document}
\thispagestyle{plain}
\myheader{Handout for Lecture 12 \\ \normalsize{Good Estimators, Sample Mean Distribution, and Confidence Intervals}}
\rhead{Class Handout for Lecture 12}

\fbox{\begin{minipage}{\textwidth}
\underline{Good Estimators} 

Denote $\hat{\theta}$ as an estimator for the population parameter $\theta$. Some desirable properties for an estimator \vspace{-1em}
\begin{itemize}
\normalsize
\item \textit{Unbiasedness}: $ E(\hat{\theta}) = \theta $
\item \textit{Efficiency}: lower variance is better
\item \textit{Consistency}: as the sample size becomes infinitely large,  $\hat{\theta} \rightarrow \theta$ 
\end{itemize}
\end{minipage}} \\~\\

\textit{Question 1:} If some sample estimator $\hat{\theta}$ is an unbiased estimator for the true population parameter $\theta$ i.e. $E(\hat{\theta}) = \theta$. This implies that: \vspace{-1em}
\begin{itemize}
\item[$\square$] $\hat{\theta}=\theta$ in all samples.
\item[$\square$] If we take repeated samples, average of $\hat{\theta}$ is equal to $\theta$ \\
\end{itemize}

\textit{Question 2}: We are choosing between two estimators $\hat{\theta}_1$ and $\hat{\theta}_2$, both of which are unbiased i.e. $E(\hat{\theta}_1)=\mu$ and $E(\hat{\theta}_2)=\mu$. But the variance of $\hat{\theta}_1$ is lower than that of $\hat{\theta}_2$ i.e. $Var(\hat{\theta}_1)<Var(\hat{\theta}_2)$.  Which of the following is true? \vspace{-1em}
\begin{itemize}
\item[$\square$] We are indifferent between the two estimators.
\item[$\square$] We prefer $\hat{\theta}_1$ over $\hat{\theta}_2$.
\item[$\square$] We prefer $\hat{\theta}_2$ over $\hat{\theta}_1$.
\item[$\square$] We need more information to reach any conclusion. \\~\\
\end{itemize}

\newpage
\fbox{\begin{minipage}{\textwidth}
\underline{Sample Mean Distribution} 

Let $X_1,X_2,...,X_n$ denote independent random draws (random sample) from a population with mean $\mu$ and variance $\sigma^2$. Then the sample mean 
$ \bar{X} = \frac{1}{n} \sum_{i=1}^n X_i $ is a random variable with:
$$E(\bar{X}) = \mu \quad \quad  Var(\bar{X}) = \frac{\sigma^2}{n} $$ 
In addition, the distribution of the sample mean is \textit{normal} if \textit{either} of the following is true: \vspace{-1em}
\begin{itemize}
  \item The underlying population is normal
  \item The sample size is large, say $n\geq 100$ 
\end{itemize}
Given the variance of the sample mean as \(\sigma_{\bar{X}}^2 = \frac{\sigma^2}{n} \), its standard deviation, commonly referred to as the \textit{standard error}, is \( \sigma_{\bar{X}} = \frac{\sigma}{\sqrt{n}} \). 
\end{minipage}} \\

\textit{Question 3}: If the average of hourly wages in the population is \( \mu = \$30 \) and the variance of hourly wages is \( \sigma^2 = 16 \). Then what is the expected value, variance, standard error, and distribution of the sample mean estimator for a sample size of 400? \\ \vspace{2cm}

\textit{Question 4}: You are interested in the average starting salary of CSUF graduates and are considering taking a random sample of 120 students. I advise you to take as large of a sample as feasible. This is sound advice because taking an even larger sample would ensure that \vspace{-1em}
\begin{itemize}
\item[$\square$] The sample average $\bar{x}=\mu$
\item[$\square$] The sample average $\bar{x}$ is drawn from a normal distribution
\item[$\square$] The sample average $\bar{x}$ is drawn from a distribution with lower variance 
\end{itemize}
\textit{Note}: I am using $\bar{x}$ to denote a realization of $\bar{X}$. \\

\textit{Question 5}: Can you explain intuitively why the variance of the sample mean increases with $\sigma^2$ and decreases with $n$?

\newpage
\fbox{\begin{minipage}{\textwidth}
\underline{Confidence Intervals}\\

Let $z_{\alpha/2}$ be the $z$-value that leaves area $\alpha/2$ in the upper tail of the normal distribution. \\
Then $1-\alpha$ confidence interval is given by 
$$ \bar{x} \pm  \underbrace{z_{\alpha/2}  \frac{\sigma}{\sqrt{n}}}_{\text{Margin of Error}} $$
\end{minipage}} \\

\textit{Question 6}: Continuing with Question 3, say you took a sample of 400 individuals and found the average hourly wages in your sample of $\bar{x} = 26$. Create a 95\% confidence interval for the true population mean. 
\end{document}