\documentclass{./../handout}

\fancyhf{}
\fancyheadoffset{1cm}
\lhead{ECON 340: Economic Research Methods}
\rhead{Div Bhagia}
\rfoot{\center \thepage}

\begin{document}
\thispagestyle{plain}
\begin{center}
  \framebox{
    \vbox{
    \vspace{0.1em}
     \Large \sc Handout for Lecture 2 \\ \vspace{0.1em}
     \large Empirical Distribution and Measures of Central Tendency \\
     \vspace{0.5em}
      \small ECON 340: Economic Research Methods \hfill Instructor: Div Bhagia
      }
    }
\end{center}	

\begin{enumerate}
\item You rolled a six-sided die 100 times and noted down how many times each of the six outcomes were realized. Fill in the rest of the table below: \\~\\
\setstretch{1.5}
\begin{tabular}{c|c|c|c}
Outcome & Count ($n_k$) & Relative frequency ($f_k$) & Cumulative frequency ($F_k$) \\
\hline
1 &  18 & & \\
\hline
2 & 18 & & \\
\hline
3 & 12 & & \\
\hline
4 & 16 & & \\
\hline
5 & 21 & & \\
\hline
6 & 15 & & \\
\hline
\textbf{Total} & \textbf{100} & \textbf{1} & \\
\hline
\end{tabular}
\vspace{0.5em}

Note that $$ f_k = \frac{n_k}{n} = \frac{\text{observations in category $k$}}{\text{total observations}} $$

\begin{enumerate}
	\item How many times did you get a die face with a value of at most 3? \\ 
	\item Are the proportions close to what you would have predicted? \\~\\
\end{enumerate}

\item Find the mean and median for: $ 3, 4, 1, 6, 8 $
\vspace{5em}
\item Amongst the mean and the median, which one is more affected by outliers? Explain. \\~\\
\item We asked a sample of 10 individuals whether they like icecream or not. We then create a variable $X$ that takes value 1 if the individual likes icecream, and 0 otherwise.  Here is the data we collected:
$$ 1, 1, 0, 0, 0, 1, 0, 1, 1, 1 $$

\begin{enumerate}
\item How many individuals like icecream in our sample? \\~\\
\item What proportion of individuals like icecream in our sample? \\~\\
\item Use the frequency distribution table and the following formula to calculate the mean of $X$. 
$$ \bar{X} = \frac{\sum_{k=1}^K n_k X_k}{n} = \sum_{k=1}^K f_k X_k $$
\end{enumerate}

\newpage
\item We have the following data on shoe sizes ($X_i$) for four individuals. 
$$ X = \{8, 6, 6, 8 \}$$
\begin{enumerate}
\item Calculate the mean:
$$ \mu = \frac {\sum_{i=1}^N X_i}{N} = \hspace{8cm} $$ \\
\item Calculate the weighted mean with weights $ w=  \{1, 1, 1, 1 \}$.
$$ \mu_{Weighted} = \frac {\sum_{i=1}^N w_i X_i}{\sum_{i=1}^N w_i } = \hspace{8cm} $$ \\
\item Calculate the weighted mean with weights $ w=  \{1, 2, 2, 1 \}$.
$$ \mu_{Weighted} = \frac {\sum_{i=1}^N w_i X_i}{\sum_{i=1}^N w_i } = \hspace{8cm} $$ \\
\item Calculate the weighted mean with weights $ w=  \{0.5, 0, 0, 0.5 \}$.
$$ \mu_{Weighted} = \frac {\sum_{i=1}^N w_i X_i}{\sum_{i=1}^N w_i } = \hspace{8cm} $$ 
\end{enumerate}
\vspace{4em}
\end{enumerate}

\end{document}