\documentclass{./../Lectures/div_teaching_slides}
\tikzset{
    -Latex,auto,node distance =1 cm and 1 cm,semithick,
    state/.style ={ellipse, draw, minimum width = 0.7 cm},
    point/.style = {circle, draw, inner sep=0.04cm,fill,node contents={}},
    bidirected/.style={Latex-Latex,dashed},
    el/.style = {inner sep=2pt, align=left, sloped}
}

\begin{document}
\title{ECON 340 \\ Economics Research Methods}
\author{Div Bhagia}
\date{Research Project Interim Submission Guidelines}

\begin{frame}[noframenumbering, plain]
\maketitle
\end{frame}




%%%%%%%%%%%% Submission 1: Overview
% \begin{frame}{Interim Submission: Overview}
% \begin{witemize}
%     \item[(a)] describe your research question and the data you will use 
%     \begin{itemize}
%       \item highlight why your question is interesting and important
%       \item mention your data sources
%       \item identify the variables you will use
%     \end{itemize}
%     \item[(b)] perform a basic analysis of your data in R
%   \end{itemize}
%   \item You need to describe your research question and the data you will use
%   \item Highlight why your question is interesting and important
%   \item Mention the source of your data
%   \item Highlight the variables you will use
% \end{witemize}
% \end{frame}

%%%%%%%%%%%% Research Question
\begin{frame}{Research Question}
\begin{witemize}
\item Your question should concern the impact of one variable on another 
\item Select a primary independent variable, a dependent variable, and at least two control variables.
\end{witemize}
\begin{center}
  \begin{tikzpicture}
  \node (1) at (0,0) {Control Variables};
  \node (2) at (-4,2) {Independent Variable};
  \node (3) at (4,2) {Dependent Variable};
  \path (2) edge  (3);
  \path[bidirected] (1) edge[bend left=20] (2);
  \path[bidirected] (3) edge[bend left=20] (1);
  \end{tikzpicture}	
  \end{center} 
\end{frame}

%%%%%%%%%%%% Data
\begin{frame}{Data}
\begin{witemize}
  \item You can pick a dataset from the Econ340 Datasets Dropbox folder or use an external dataset
  \item There is a list of all the datasets in the folder; each folder contains the codebook (description of variables)
  \item  If you need help with formatting an external dataset, email me at least a week in advance
  \item Your dataset should meet the following criteria:\\
  \begin{itemize}
  \item Have at least 100 observations
  \item Contain all the variables you plan to use
  \end{itemize}
\end{witemize}
\end{frame}

%%%%%%%%%%%% Overview
\begin{frame}{Interim Submission: Overview}
  \textbf{Goals}:  \vskip0.5em
  \begin{witemize}
      \item[(a)] Describe your research question and data 
      \item[(b)] Perform a basic analysis of your data in R   \\~\\
  \end{witemize}
  The submission will serve as a foundation for the “Introduction” and “Descriptive Analysis” sections of your final paper. \\~\\
  Refer to the example submission and it’s corresponding R code to guide you.
  \end{frame}

%%%%%%%%%%%% Tasks
\begin{frame}{Tasks}
\begin{wenumerate}
  \item Complete the cover sheet.
  \item Specify and motivate your research question.
  \item Describe your data and variables. All variables should be either continuous or binary.
  \item Present summary statistics for all your variables and comment.
\end{wenumerate}
\end{frame}

%%%%%%%%%%%% Tasks
\begin{frame}{Tasks}
\begin{witemize}
\item[5.] Present a graph showing the relationship between your dependent and primary independent variable \\ \vskip0.5em
\begin{itemize}
\item Scatter plot if both variables are continuous
\item Bar plot if one variable is binary \\ \vskip0.5em
\end{itemize}
Also, report the correlation between the two variables. Summarize what you learn. 
\item[6.] Present the correlation between all your variables in one or more tables and discuss. Comment on the direction of bias from not accounting for control variables. 
\end{witemize}
\end{frame}

%%%%%%%%%%%% Submission
\begin{frame}{Submission}
\begin{witemize}
  \item Prepare your submission in Microsoft Word and submit it in PDF format, along with R code and data (if changed significantly) on Canvas.
  \item Export figures from R and include them in your document. 
  \item For tables, you can copy from the output and format them as a table in Word (use a fixed-width font like Courier New) or take a screenshot.
  \item All tables and figures should have a name and number. 
  \item The write-up should be in essay format, no bullet points.
\end{witemize}
\end{frame}

\end{document}